---
title: Analysis of Violence Ratings and Bechdel Test Scores for Oscars Nominated Films
  from 2020 to 2026
author: "Esther Lim, Ethan Sanders, Kamila Kwiatkowska"
date: "2026-04-19"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

# Research Question

### Among the Oscar Academy Award Best Picture shortlisted films from 2020 to 2026, is there a significant relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets?

# Hypotheses

### Hypothesis 1 (H1)

There is a statistically detectable relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets among Oscar Best Picture shortlisted films (2020-2026).

### Hypothesis 2 (H2)

There is a relationship between a film being female-directed and non-female-directed and its odds of passing all three Bechdel Test criteria among Oscar Best Picture shortlisted films (2020-2026).

### Hypothesis 3 (H3)

There is a relationship between a film’s worldwide box office revenue and the IMDb Violence and Gore rating among Oscar Best Picture shortlisted films (2020-2026).

# Summarised Findings

### Hypothesis 1 (H1)

There is a statistically detectable relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets among Oscar Best Picture shortlisted films (2020-2026).

### Hypothesis 2 (H2)

There is a statistically detectable relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets among Oscar Best Picture shortlisted films (2020-2026).

### Hypothesis 3 (H3)

There is a statistically detectable relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets among Oscar Best Picture shortlisted films (2020-2026).

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(readr)
library(ggplot2)
library(paletteer)

options(scipen = 999)

film_data <- read.csv("data/violence-bechdel-data.csv")
  skip_empty_rows = TRUE
  stringsAsFactors = TRUE
  
# Hypothesis 1 Data Setup
  
# Hypothesis 2 Data Setup

# Hypothesis 3 Data Setup
  
```

## H1 Analysis

### Chi-square Test of Independence (H1)

```{r Chi-square Test H1}

violence_levels <- c(
  'None',
  'Mild',
  'Moderate',
  'Severe'
)

film_data <- film_data %>%
  mutate(
    violence = violence %>%
      ordered(
        levels = violence_levels
      )
  )

h1chisq_tbl <- table(film_data$violence, film_data$bechdel)

print("2x4 Contingency Table:")
print(h1chisq_tbl)

h1chisq_results <- chisq.test(h1chisq_tbl)

print("H1 Chi-square Test Results:")
print(h1chisq_results)

```

#### Chi-square Test of Independence (H1) Results

p-value = 0.097. Because of the sample size, a Fisher's Exact Test is recommended.

### Fisher's Exact Test (H1)

```{r Fishers Exact Test H1}

h1fsh_tbl <- table(film_data$violence, film_data$bechdel)
print("4x4 Contingency Table:")
print(h1fsh_tbl)

h1fsh_results <- fisher.test(table(film_data$violence, film_data$bechdel))
print("H1 Fisher's Exact Test Results:")
print(h1fsh_results)

```

#### Fisher's Exact Test (H1) Results

p-value = 0.087.

### OLS Linear Model (H1)

```{r OLS H1}

h1ols <- lm(film_data$bechdel ~ film_data$violence, film_data = ols_data) 

summary(h1ols)

```

#### OLS (H1) Results

p-value = 0.123.

## H1 Data Visualisation

### Scatterplot (H1)

```{r Scatterplot H1}

film_data %>%
  ggplot(
    aes(x = violence, y = bechdel)
  ) +
  geom_point(color = "#00496FFF") +
  labs(title = "Scatterplot for Hypothesis 1",
       x = "IMDb Violence and Gore score",
       y = "Bechdel Test criteria passed")

```

### Scatterplot with Jitter (H1)

```{r Scatterplot Jitter H1}

film_data %>%
  ggplot(
    aes(x = violence, y = bechdel)
  ) +
  geom_jitter(color = "#ED8B00FF") +
  labs(title = "Scatterplot for Hypothesis 1",
       x = "IMDb Violence and Gore score",
       y = "Bechdel Test criteria passed")

```

### Scatterplot with Regression (H1)

```{r Scatterplot H1 w Regression}

film_data %>%
  ggplot(
    aes(x = violence_coded, y = bechdel)
  ) +
  geom_jitter(color = "#ED8B00FF") +
geom_smooth(method = "lm", se = FALSE, col = "#DD4124FF") +
  labs(title = "Scatterplot with Regression for Hypothesis 1",
       x = "IMDb Violence and Gore score",
       y = "Bechdel Test criteria passed")

```

### Spearman Rho Test (H1)

```{r Spearman Rho Test H1}
  
corr_results <- cor.test(x = film_data$violence_coded, y = film_data$bechdel, method = "spearman")
print(corr_results)

```

#### Spearman Rho Test Results (H1)

p-value = 0.548

### Violin (H1)

```{r Violin H1}
  
film_data %>%
  ggplot(
    aes(x = violence, y = bechdel)
  ) +
  geom_violin(color = "#ED8B00FF", fill = "#EDD746FF")

```

### Boxplot (H1)

```{r Boxplot H1}

film_data %>%
  ggplot(
    aes(x = violence, y = bechdel)
  ) +
  geom_boxplot(color = "#ED8B00FF", fill = "#EDD746FF")

```
---

## Hypothesis 2 (H2)

There is a relationship between a film being female-directed and non-female-directed and its odds of passing all three Bechdel Test criteria among Oscar Best Picture shortlisted films (2020-2026).

## H2 Analysis

### Fisher's Exact Test (H2)

```{r}

odds_test <- film_data %>%
  mutate(director_binary = factor(
      ifelse(gen_dir == "female", "Female", "Non-Female"),
      levels = c("Non-Female", "Female")
      ),
    bechdel_binary = factor(
      ifelse(bechdel == 3, "Pass", "Fail"),
      levels = c("Fail", "Pass")
    )
  )

odds_table <- table(odds_test$director_binary, odds_test$bechdel_binary)

print("2x2 Contingency Table:")
print(odds_table)

odds_results <- fisher.test(odds_table)

print("Odds Ratio Test Results:")
print(odds_results)

```

#### Fisher's Exact Test (H2) Results

A Fisher’s exact test was conducted to examine the relationship between director gender and passing the Bechdel test. While female-directed films had over five times higher odds of passing the test compared to non-female-directed films, this difference was not statistically significant, OR = 5.53, with a 95% confidence interval [0.67, 260.50], p = .138.

### Chi-square Test (H2)

```{r chisquare H2}

h2chisq_tbl <- table(film_data$gen_dir, film_data$bechdel)
print("Observed Frequencies:")
print(h2chisq_tbl)

h2chisq_results <- chisq.test(h2chisq_tbl)

print("Expected Frequencies (Must mostly be >= 5):")
print(h2chisq_results$expected)

print("Chi-Square Test Results:")
print(h2chisq_results)

ggplot(film_data, aes(x = bechdel)) +
  geom_histogram(binwidth = 1, fill = "#0F85A0FF", color = "#89AAC6FF") +
  facet_wrap(~ gen_dir) +
  theme_minimal() +
  labs(title = "Distribution of Bechdel Scores by Director Gender (Normality Check)")

```

#### Chi-square Test (H2) Results

A chi-square test was conducted to examine if there is a relationship between director gender and Bechdel Test score. As the number of female directors in the dataset were too few, the calculated chi-square value was inaccurate. Furthermore, this relationship was not statistically significant, chi-square = 4.8575, df = 3, p = .183.

### Independent Samples t-test (H2)

```{r t test H2}

h2tt_results <- t.test(bechdel ~ gen_dir, data = film_data)

print("Independent Samples t-test Results:")
print(h2tt_results)

```

#### Independent Samples t-test (H2) Results

A independent samples t-test was thus conducted to examine the difference between the mean average  score on the Bechdel Test between genders of directors. This assumes that the variable for the Bechdel Test is continuous. The result showed that there was no statistically significant difference between the Bechdel Test scores of female-directed movies and non-female-directed movies, t = 1.1415, df = 10.548, p = .279.

### Mann-Whitney U Test (H2)

```{r Mann-Whitney U Test H2}

film_data %>%
  group_by(gen_dir) %>%
  summarize(median_bechdel = median(bechdel, na.rm = TRUE))

h2mannwhit_results <- wilcox.test(bechdel ~ gen_dir, data = film_data)
print("Mann-Whitney U Test Results:")
print(h2mannwhit_results)

```

#### Mann-Whitney U Test (H2) Results

In order to account for few cell counts for female director variable, and to treat the Bechdel Test score as ordinal, a Mann-Whitney U Test was conducted to determine if there was a significant difference in Bechdel Test scores between female-directed and non-female-directed Oscar-nominated films from 2000 to 2026. The median Bechdel Test score for female-directed films (Mdn = 3) was not significantly different from non-female-directed films (Mdn = 3), U = 329.5, p = 0.147.

## H2 Data Visualisation

### Bar Plot

This bar plot shows the breakdown of Bechdel Test criteria met by each film grouped according to gender of film director (Female, Non-female).

```{r bar plot H2}

ggplot(film_data, aes(x = gen_dir, fill = factor(bechdel))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_paletteer_d("PNWColors::Bay") +
  theme_minimal() +
  labs(
    title = "Proportion of Bechdel Test Scores by Director Gender",
    x = "Director Gender",
    y = "Percentage of Films",
    fill = "Bechdel Test Score"
  )
```
### Time Trend

This time trend graph shows the proportion of films that meet all three criteria of the Bechdel Test by year, from 2020 to 2026, split by gender of film director. 

```{r time trend H2}

yearly_trends <- film_data %>%
  mutate(
    Passed = ifelse(bechdel == 3, 1, 0),
    director_binary = factor(ifelse(gen_dir == "female", "Female", "Non-Female"))
  ) %>%
  group_by(year, director_binary) %>%
  summarize(Pass_Rate = mean(Passed, na.rm = TRUE), .groups = "drop")

ggplot(yearly_trends, aes(x = year, y = Pass_Rate, color = director_binary)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = c("Non-Female" = "#0F85A0FF", "Female" = "#DD4124FF")) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  theme_minimal() +
  labs(
    title = "Bechdel Test Pass Rates Over Time (2020-2026)",
    x = "Year of Release",
    y = "Percentage of Films Passing (Score = 3)",
    color = "Director Gender"
  )

```

### Cumulative Density Plot

This cumulative density plot visualises the proportion of films over the six years across the Bechdel Test Scores, split by gender of film director.

```{r cumulative density H2}

ggplot(film_data, aes(x = bechdel, color = gen_dir)) +
  stat_ecdf(geom = "step", size = 1.2) +
  scale_color_manual(values = c("#DD4124FF", "#0F85A0FF")) +
  theme_minimal() +
  labs(
    title = "Cumulative Density of Bechdel Test Scores by Gender",
    x = "Bechdel Test Score (0 to 3)",
    y = "Cumulative Proportion",
    color = "Director Gender"
  )

```

### Mosaic Plot 

This mosaic plot visualises distribution of Bechdel Test criteria met by the whole population-Oscars Academy Award Best Picture Shortlisted Films between 2020 and 2026-split by gender of film director and proportions of each number of Bechdel Test criteria met.

```{r mosaic plot}

mosaic_table <- table(film_data$gen_dir, film_data$bechdel_coded)

mosaicplot(
  mosaic_table, 
  main = "Mosaic Plot of Director Gender vs. Bechdel Test Fail/Pass",
  color = c("#ED8B00FF", "#0F85A0FF"),
  xlab = "Director Gender", 
  ylab = "Bechdel Status (Fail / Pass)",
  border = "#00496FFF"
)

mosaic_table_more <- table(film_data$gen_dir, film_data$bechdel)

mosaicplot(
  mosaic_table_more, 
  main = "Mosaic Plot of Director Gender vs. Bechdel Test Score",
  color = c("#ED8B00FF", "#89AAC6FF"),
  xlab = "Director Gender", 
  ylab = "Bechdel Test Score",
  border = "#00496FFF"
)

```

## Hypothesis 3 (H3)

There is a relationship between a film’s worldwide box office revenue and the IMDb Violence and Gore rating among Oscar Best Picture shortlisted films (2020-2026).

## H3 Analysis

### Test Name (H3)

```{r}

# Test code here!

```

#### Test Name (H3) Results


## Data Visualisation

### Visualisation Name (H3)

``` {r}

# Plot code here!

```
