---
title: "Violence and Female Representation in Movies"
author: "Esther Lim, Ethan Sanders, Kamila Kwiatkowska"
date: "2026-04-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(readr)
library(ggplot2)
options(scipen = 999)

film_data <- read.csv("data/violence-bechdel-data.csv")
  skip_empty_rows = TRUE
  stringsAsFactors = TRUE
  
```

# Final Project Analysis

R Project and Markdown document set up

# Research Question

### Among the Oscar Academy Award Best Picture shortlisted films from 2020 to 2026, is there a significant relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets?

# Hypotheses

### Hypothesis 1 (H1)

There is a statistically detectable relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets among Oscar Best Picture shortlisted films (2020-2026).

### Hypothesis 2 (H2)

There is a relationship between a film being female-directed and non-female-directed and its odds of passing all three Bechdel Test criteria among Oscar Best Picture shortlisted films (2020-2026).

### Hypothesis 3 (H3)

There is a relationship between a film’s worldwide box office revenue and the IMDb Violence and Gore rating among Oscar Best Picture shortlisted films (2020-2026).

# Findings

## Hypothesis 1 (H1)

There is a statistically detectable relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets among Oscar Best Picture shortlisted films (2020-2026).

### Chi-square Test of Independence (H1)

```{r Chi-square Test H1}

violence_levels <- c(
  'None',
  'Mild',
  'Moderate',
  'Severe'
)

film_data <- film_data %>%
  mutate(
    violence = violence %>%
      ordered(
        levels = violence_levels
      )
  )

table(film_data$violence, film_data$bechdel)

test <- chisq.test(table(film_data$violence, film_data$bechdel))
test

```

#### Chi-square Test of Independence (H1) results

p-value = 0.097. Because of the sample size, a Fisher's Exact Test is recommended.

### Fisher's Exact Test (H1)

```{r Fishers Exact Test H1}

table(film_data$violence, film_data$bechdel)

test <- fisher.test(table(film_data$violence, film_data$bechdel))
test

```

#### Fisher's Exact Test (H1) results

p-value = 0.087.

### OLS Linear Model

```{r OLS H1}

ols <- lm(film_data$bechdel ~ film_data$violence, film_data = ols_data) 
ols
summary(ols)

```

#### OLS (H1) results

p-value = 0.123.

### Scatterplot (H1)

```{r}

film_data %>%
  ggplot(
    aes(x = violence, y = bechdel)
  ) +
  geom_point(color = "lightgreen") +
  labs(title = "Scatterplot for Hypothesis 1",
       x = "IMDb Violence and Gore score",
       y = "Bechdel Test criteria passed")

```

### Scatterplot with Jitter (H1)

```{r}

film_data %>%
  ggplot(
    aes(x = violence, y = bechdel)
  ) +
  geom_jitter(color = "darkgreen") +
  labs(title = "Scatterplot for Hypothesis 1",
       x = "IMDb Violence and Gore score",
       y = "Bechdel Test criteria passed")

```

### Scatterplot with Regression (H1)

```{r Scatterplot H1 w Regression}

film_data %>%
  ggplot(
    aes(x = violence_coded, y = bechdel)
  ) +
  geom_jitter(color = "darkgreen") +
geom_smooth(method = "lm", se = FALSE, col = "lightgreen") +
  labs(title = "Scatterplot with Regression for Hypothesis 1",
       x = "IMDb Violence and Gore score",
       y = "Bechdel Test criteria passed")

```

### Spearman Rho Test (H1)

```{r Spearman Rho Test H1}
  
corr <- cor.test(x = film_data$violence_coded, y = film_data$bechdel, method = "spearman")
corr

```

#### Spearman Rho Test Results (H1)

p-value = 0.548

### Violin (H1)

```{r Violin H1}
  
film_data %>%
  ggplot(
    aes(x = violence, y = bechdel)
  ) +
  geom_violin(color = "darkgreen", fill = "lightgreen")
```

### Boxplot (H1)

```{r}

film_data %>%
  ggplot(
    aes(x = violence, y = bechdel)
  ) +
  geom_boxplot(fill = "lightgreen", color = "darkgreen")

```

## Hypothesis 2 (H2)

There is a relationship between a film being female-directed and non-female-directed and its odds of passing all three Bechdel Test criteria among Oscar Best Picture shortlisted films (2020-2026).

## Hypothesis 3 (H3)

There is a relationship between a film’s worldwide box office revenue and the IMDb Violence and Gore rating among Oscar Best Picture shortlisted films (2020-2026).