---
title: Analysis of Violence Ratings and Bechdel Test Scores for Oscars Nominated Films
  from 2020 to 2026
author: "Esther Lim, Ethan Sanders, Kamila Kwiatkowska"
date: "2026-04-19"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

# Research Question

### Among the Oscar Academy Award Best Picture shortlisted films from 2020 to 2026, is there a significant relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets?

# Hypotheses

### Hypothesis 1 (H1)

There is a statistically detectable relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets among Oscar Best Picture shortlisted films (2020-2026).

### Hypothesis 2 (H2)

There is a relationship between a film being female-directed and non-female-directed and its odds of passing all three Bechdel Test criteria among Oscar Best Picture shortlisted films (2020-2026).

### Hypothesis 3 (H3)

There is a relationship between a film’s worldwide box office revenue and the IMDb Violence and Gore rating among Oscar Best Picture shortlisted films (2020-2026).

# Summarised Findings

### Hypothesis 1 (H1)

No statistically significant relationship detected between violence rating and Bechdel score. Notable: cubic contrast in OLS was significant (p = .041), suggesting possible non-linear pattern worth investigating.

### Hypothesis 2 (H2)

Female-directed films showed 5.5× higher odds of passing the Bechdel test, but the result was non-significant, almost certainly due to very small female-director cell count (n = 9).

### Hypothesis 3 (H3)

Revenue distributions were broadly similar across all four violence rating groups. Descriptively, "Severe" films had the highest median revenue ($117M), but this difference was not significant.

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

my_packages <- c("tidyverse", "ggplot2", "paletteer","scales","dunn.test")

for (pkg in my_packages) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg, dependencies = TRUE)
    library(pkg, character.only = TRUE)
  }
}

options(scipen = 999)

film_data <- read.csv("data/violence-bechdel-data.csv")
  skip_empty_rows = TRUE
  stringsAsFactors = TRUE
  
```

## H1 Analysis

### Chi-square Test of Independence (H1)

```{r Chi-square Test H1}

violence_levels <- c('None', 'Mild', 'Moderate', 'Severe')

film_data <- film_data %>%
  mutate(
    violence = violence %>%
      ordered(levels = violence_levels)
  )

h1chisq_tbl <- table(film_data$violence, film_data$bechdel)

print("2x4 Contingency Table:")
print(h1chisq_tbl)

h1chisq_results <- chisq.test(h1chisq_tbl)

print("H1 Chi-square Test Results:")
print(h1chisq_results)

```

#### Chi-square Test of Independence (H1) Results

A chi-square test of independence was conducted to examine the relationship between IMDb Violence and Gore rating (None, Mild, Moderate, Severe) and the number of Bechdel Test criteria met (0–3) among Oscar Best Picture shortlisted films from 2020 to 2026. As several expected cell counts fell below 5, the result should be interpreted with caution, as the chi-square approximation may be unreliable. The result was not statistically significant, X2 (df = 9, n = 67) = 14.79, p = .097. A Fisher's exact test is therefore recommended as a more appropriate alternative given the sample size.

### Fisher's Exact Test (H1)

```{r Fishers Exact Test H1}

h1fsh_tbl <- table(film_data$violence, film_data$bechdel)
print("4x4 Contingency Table:")
print(h1fsh_tbl)

h1fsh_results <- fisher.test(table(film_data$violence, film_data$bechdel))
print("H1 Fisher's Exact Test Results:")
print(h1fsh_results)

```

#### Fisher's Exact Test (H1) Results

A Fisher's exact test was conducted to examine the association between IMDb Violence and Gore rating and Bechdel Test score, as the small sample size rendered the chi-square approximation unreliable. The test was not statistically significant, p = .087. The null hypothesis is therefore retained, as the data provide no evidence of an association between a film's violence rating and its Bechdel Test score among Oscar Best Picture shortlisted films (2020–2026).

### OLS Linear Model (H1)

```{r OLS H1}

h1ols <- lm(film_data$bechdel ~ film_data$violence, film_data = ols_data) 

summary(h1ols)

```

#### OLS (H1) Results

An ordinary least squares regression was conducted to examine whether IMDb Violence and Gore rating predicted the number of Bechdel Test criteria met. The overall model was not statistically significant, F(3, 63) = 2.00, p = .123, R2 = .087, adjusted R2 = .044. The null hypothesis is therefore retained. However, it is worth noting that the cubic polynomial contrast was statistically significant, b = −0.459, SE = 0.220, t(63) = −2.08, p = .041, which may suggest a non-linear component to the relationship that warrants further investigation with a larger sample.

### Spearman Rho Test (H1)

```{r Spearman Rho Test H1}
  
h1corr_results <- cor.test(x = film_data$violence_coded, y = film_data$bechdel, method = "spearman")
print(h1corr_results)

```

#### Spearman Rho Test Results (H1)

A Spearman's rank correlation was conducted to examine the monotonic relationship between IMDb Violence and Gore rating and the number of Bechdel Test criteria met. The warning regarding ties is expected when an ordinal variable with few levels is used and does not invalidate the result. The correlation was negligible and not statistically significant, rs(65) = −.075, S = 53861, p = .548. The null hypothesis is therefore retained, as the data provide no evidence of a monotonic relationship between a film's violence rating and its Bechdel Test score among Oscar Best Picture shortlisted films (2020–2026).


## H1 Data Visualisation

### Scatterplot (H1)

```{r Scatterplot H1}

film_data %>%
  ggplot(aes(x = violence, y = bechdel)) +
  geom_point(color = "#ED8B00FF") +
  
  labs(
    title = "Bechdel Scores by Violence Rating",
    x = "Violence and Gore Score",
    y = "Bechdel Test Score")

```

### Scatterplot with Jitter (H1)

```{r Scatterplot Jitter H1}

film_data %>%
  ggplot(aes(x = violence, y = bechdel)) +
  geom_jitter(width = 0.25, color = "#ED8B00FF") +
  labs(
    title = "Bechdel Scores by Violence Rating",
    x = "Violence and Gore Score",
    y = "Bechdel Test Score")

```

### Scatterplot with Regression (H1)

```{r Scatterplot H1 w Regression}

film_data %>%
  ggplot(aes(x = violence_coded, y = bechdel)) +
  geom_jitter(width = 0.25, color = "#ED8B00FF") +
  geom_smooth(
    aes(group = 1), 
    method = "lm", 
    se = FALSE, 
    col = "#00496FFF"
    ) +
  
  labs(
    title = "Bechdel Scores by Violence Rating",
    x = "Violence and Gore Score",
    y = "Bechdel Test Score")

```



### Violin (H1)

```{r Violin H1}
  
film_data %>%
  ggplot(aes(x = violence, y = bechdel)) +
  geom_violin(color = "#ED8B00FF", fill = "#EDD746FF")+
  
  labs(
    title = "Distribution of Bechdel Scores by Violence Rating",
    x = "Violence and Gore Rating",
    y = "Bechdel Test Score"
  )

```

### Boxplot (H1)

```{r Boxplot H1}

film_data %>%
  ggplot(aes(x = violence, y = bechdel)) +
  geom_boxplot(color = "#ED8B00FF", fill = "#EDD746FF")

labs(
    title = "Distribution of Bechdel Scores by Violence Rating",
    x = "Violence and Gore Rating",
    y = "Bechdel Test Score"
)

```
---

## Hypothesis 2 (H2)

There is a relationship between a film being female-directed and non-female-directed and its odds of passing all three Bechdel Test criteria among Oscar Best Picture shortlisted films (2020-2026).

## H2 Analysis

### Fisher's Exact Test (H2)

```{r fishers test H2}

h2odds_test <- film_data %>%
  mutate(director_binary = factor(
      ifelse(gen_dir == "female", "Female", "Non-Female"),
      levels = c("Non-Female", "Female")
      ),
    bechdel_binary = factor(
      ifelse(bechdel == 3, "Pass", "Fail"),
      levels = c("Fail", "Pass")
    )
  )

h2odds_table <- table(h2odds_test$director_binary, h2odds_test$bechdel_binary)

print("2x2 Contingency Table:")
print(h2odds_table)

h2odds_results <- fisher.test(h2odds_table)

print("Odds Ratio Test Results:")
print(h2odds_results)

```

#### Fisher's Exact Test (H2) Results

A Fisher’s exact test was conducted to examine the relationship between director gender and passing the Bechdel test. While female-directed films had over five times higher odds of passing the test compared to non-female-directed films, this difference was not statistically significant, OR = 5.53, with a 95% confidence interval [0.67, 260.50], p = .138.

### Chi-square Test (H2)

```{r chisquare H2}

h2chisq_tbl <- table(film_data$gen_dir, film_data$bechdel)
print("Observed Frequencies:")
print(h2chisq_tbl)

h2chisq_results <- chisq.test(h2chisq_tbl)

print("Expected Frequencies (Must mostly be >= 5):")
print(h2chisq_results$expected)

print("Chi-Square Test Results:")
print(h2chisq_results)

ggplot(film_data, aes(x = bechdel)) +
  geom_histogram(binwidth = 1, fill = "#0F85A0FF", color = "#89AAC6FF") +
  facet_wrap(~ gen_dir) +
  theme_minimal() +
  labs(
    title = "Distribution of Bechdel Scores by Director's Gender \n (Normality Check)",
    x = "Bechdel Test Score",
    y = "Count"
    )

```

#### Chi-square Test (H2) Results

A chi-square test was conducted to examine if there is a relationship between director gender and Bechdel Test score. As the number of female directors in the dataset were too few, the calculated chi-square value was inaccurate. Furthermore, this relationship was not statistically significant, chi-square = 4.8575, df = 3, p = .183.

### Independent Samples t-test (H2)

```{r t test H2}

h2tt_results <- t.test(bechdel ~ gen_dir, data = film_data)

print("Independent Samples t-test Results:")
print(h2tt_results)

```

#### Independent Samples t-test (H2) Results

A independent samples t-test was thus conducted to examine the difference between the mean average  score on the Bechdel Test between genders of directors. This assumes that the variable for the Bechdel Test is continuous. The result showed that there was no statistically significant difference between the Bechdel Test scores of female-directed movies and non-female-directed movies, t = 1.1415, df = 10.548, p = .279.

### Mann-Whitney U Test (H2)

```{r Mann-Whitney U Test H2}

film_data %>%
  group_by(gen_dir) %>%
  summarize(median_bechdel = median(bechdel, na.rm = TRUE))

h2mannwhit_results <- wilcox.test(bechdel ~ gen_dir, data = film_data)
print("Mann-Whitney U Test Results:")
print(h2mannwhit_results)

```

#### Mann-Whitney U Test (H2) Results

In order to account for few cell counts for female director variable, and to treat the Bechdel Test score as ordinal, a Mann-Whitney U Test was conducted to determine if there was a significant difference in Bechdel Test scores between female-directed and non-female-directed Oscar-nominated films from 2000 to 2026. The median Bechdel Test score for female-directed films (Mdn = 3) was not significantly different from non-female-directed films (Mdn = 3), U = 329.5, p = 0.147.

## H2 Data Visualisation

### Bar Plot (H2)

This bar plot shows the breakdown of Bechdel Test criteria met by each film grouped according to gender of film director (Female, Non-female).

```{r bar plot H2}

ggplot(film_data, aes(x = gen_dir, fill = factor(bechdel))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_paletteer_d("PNWColors::Bay") +
  theme_minimal() +
  labs(
    title = "Proportion of Bechdel Test Scores by Director Gender",
    x = "Director's Gender",
    y = "Percentage of Films",
    fill = "Bechdel Test Score"
  )
```
### Time Trend (H2)

This time trend graph shows the proportion of films that meet all three criteria of the Bechdel Test by year, from 2020 to 2026, split by gender of film director. 

```{r time trend H2}

yearly_trends <- film_data %>%
  mutate(
    Passed = ifelse(bechdel == 3, 1, 0),
    director_binary = factor(ifelse(gen_dir == "female", "Female", "Non-Female"))
  ) %>%
  group_by(year, director_binary) %>%
  summarize(Pass_Rate = mean(Passed, na.rm = TRUE), .groups = "drop")

ggplot(yearly_trends, aes(x = year, y = Pass_Rate, color = director_binary)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = c("Non-Female" = "#0F85A0FF", "Female" = "#DD4124FF")) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  theme_minimal() +
  labs(
    title = "Bechdel Test Pass Rates Over Time (2020-2026)",
    x = "Year of Release",
    y = "Percentage of Films Passing (Score = 3)",
    color = "Director Gender"
  )

```

### Cumulative Density Plot (H2)

This cumulative density plot visualises the proportion of films over the six years across the Bechdel Test Scores, split by gender of film director.

```{r cumulative density H2}

ggplot(film_data, aes(x = bechdel, color = gen_dir)) +
  stat_ecdf(geom = "step", size = 1.2) +
  scale_color_manual(values = c("#DD4124FF", "#0F85A0FF")) +
  theme_minimal() +
  labs(
    title = "Cumulative Density of Bechdel Test Scores by Gender",
    x = "Bechdel Test Score (0 to 3)",
    y = "Cumulative Proportion",
    color = "Director Gender"
  )

```

### Mosaic Plot (H2)

This mosaic plot visualises distribution of Bechdel Test criteria met by the whole population-Oscars Academy Award Best Picture Shortlisted Films between 2020 and 2026-split by gender of film director and proportions of each number of Bechdel Test criteria met.

```{r mosaic plot}

mosaic_table <- table(film_data$gen_dir, film_data$bechdel_coded)

mosaicplot(
  mosaic_table, 
  main = "Mosaic Plot of Director Gender \nvs. Bechdel Test Fail/Pass",
  color = c("#ED8B00FF", "#0F85A0FF"),
  xlab = "Director Gender", 
  ylab = "Bechdel Status (Fail / Pass)",
  border = "#00496FFF"
)

mosaic_table_more <- table(film_data$gen_dir, film_data$bechdel)

mosaicplot(
  mosaic_table_more, 
  main = "Mosaic Plot of Director's Gender \nvs. Bechdel Test Score",
  color = c("#ED8B00FF", "#89AAC6FF"),
  xlab = "Director's Gender", 
  ylab = "Bechdel Test Score",
  border = "#00496FFF"
)

```

## Hypothesis 3 (H3)

There is a relationship between a film’s worldwide box office revenue and the IMDb Violence and Gore rating among Oscar Best Picture shortlisted films (2020-2026).

## H3 Analysis

##Descriptive Statistics
```{r data preparation}

violence_levels <- c('None','Mild','Moderate','Severe')

film_data <- film_data %>%
  mutate(
    box_office = box_office %>%
      as.numeric(),

    log_box_office = log10(box_office),

    violence = factor(
      violence, 
      levels = violence_levels, 
      ordered = TRUE
      ),

    violence_coded = as.numeric(factor(
      violence, 
      levels = violence_levels, 
      ordered = TRUE))

) %>%

  filter(!is.na(box_office))
  
```

```{r descriptive statistics}
film_data %>%
  group_by(violence) %>%
  summarize(
    n = n(),
    Mdn = median(box_office, na.rm = TRUE),
    M = mean(box_office, na.rm = TRUE),
    SD = sd(box_office, na.rm = TRUE),
    Mdn_log = median(log_box_office, na.rm = TRUE),
    M_log = mean(log_box_office, na.rm = TRUE)
  )%>%
  
  mutate(
    Mdn = scales::dollar(Mdn, accuracy = 1),
    M = scales::dollar(M, accuracy = 1),
    SD = scales::dollar(SD, accuracy = 1),
    Mdn_log = round(Mdn_log, 2),
    M_log = round(M_log, 2)
  )

```

### Kruskal-Wallis Test (H3)

```{r kruskal}

h3krus_results <- kruskal.test(
  box_office ~ violence, 
  data = film_data
)

```

#### Kruskal-Wallis Test (H3) Results

A Kruskal-Wallis test was conducted to examine whether worldwide box office revenue differed across violence rating groups. The test was not statistically significant, H(3) = 1.53, p = .676. The null hypothesis is therefore retained, as the data provide no evidence of a relationship between a film's violence rating and its worldwide box office revenue among Oscar Best Picture shortlisted films (2020–2026).

### Dunn Post-Hoc Test (H3)

```{r dunn}

h3dunn_results <- dunn.test(
  x = film_data$box_office,
  g = film_data$violence,
  method = 'bonferroni'
)

```

#### Dunn Post-Hoc Test (H3) Results

As the Kruskal-Wallis test was not statistically significant, Dunn's pairwise post-hoc comparisons are reported for completeness only and should not be interpreted inferentially. All Bonferroni-corrected pairwise comparisons were non-significant, with p-values ranging from .705 to 1.000, confirming that no specific pair of violence rating groups differed significantly in box office revenue.

### Spearman Correlation (H3)

```{r spearman}

h3sp_results <- cor.test(
  ~ as.integer(violence) + box_office,
  data   = film_data,
  method = 'spearman'
)

```
#### Spearman Correlation (H3) Results

A Spearman correlation was conducted to examine the monotonic relationship between violence rating and worldwide box office revenue. The warning regarding ties is expected when an ordinal variable with few levels is used and does not invalidate the result. The correlation was weak and not statistically significant, rho = .119, S = 36711, p = .353, consistent with the Kruskal-Wallis finding. The null hypothesis is therefore retained.

## Data Visualisation

### Scatter Plot (H3)

``` {r scatter plot}

film_data %>%
  ggplot(aes(x = violence, y = log_box_office)) +
  geom_jitter(width = 0.25, color = "#ED8B00FF") +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE, col = "#00496FFF") +

  scale_y_continuous(
    name = expression(paste("Worldwide Box Office ($, ", log[10], ")")),
    breaks = seq(5, 10, by = 1),
  ) +
    
    labs(
    x     = 'Violence and Gore Rating',
    title = 'Worldwide Box Office Revenue by Violence Rating'
  )
```

### Violon Plot (H3)

``` {r violin plot}

film_data %>%
  ggplot(aes(x = violence, y = log_box_office)) +

  geom_violin(color = "#ED8B00FF", fill = "#EDD746FF") +
  
  scale_y_continuous(
    name = expression(paste("Worldwide Box Office ", log[10], " ($)")),
    breaks = seq(3, 10, by = 1),
    labels = scales::math_format(10^.x),
    limits = c(3, 10)
  ) +
    
  labs(
    x     = 'Violence and Gore Rating',
    title = 'Distribution of Box Office Revenue by Violence Rating'
  )

```

### Box Plot (H3)
``` {r box plot}

film_data %>%
  ggplot(aes(x = violence, y = log_box_office)) +

  geom_boxplot(color = "#ED8B00FF", fill = "#EDD746FF") +
  
  
  scale_y_continuous(
    name = expression(paste("Worldwide Box Office ", log[10], " ($)")),
    breaks = seq(3, 10, by = 1),
    labels = scales::math_format(10^.x),
    limits = c(3, 10)
  ) +
    
  labs(
    x = 'Violence and Gore Rating',
    title = 'Worldwide Box Office Revenue by Violence Rating'
  )

```